---
title: 'Problem Set 2: Omitted Variable Bias and Fixed Effects'
author: "Claire Duquennois"
output:
  pdf_document: default
  html_document:
    number_sections: yes
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
***NAME:Nachu Lakshmanan***


**Empirical Analysis using Data from Washington (2008, AER)**


This exercise, like PS1, also uses data from Ebonya Washington's paper, "Female Socialization: How Daughters Affect their Legislator Father's voting on Women's Issues," published in the *American Economic Review* in 2008. This paper studies whether having a daughter affects legislator's voting on women's issues. 


**Submission instructions:**

1) Knit your assignment in PDF.

2) **Assignment shell extra credit:** You will receive a little extra credit if your answers line up correctly with the answer positions of the template on gradescope. For this to work 

- **Work by putting your answers directly into this R markdown file that is pre-formatted for you!**

- Make sure you have ONE question and answer per page (this is how the template is structured and  allows gradescope to easily find your answers), unless the question indicated that the answer will require two pages because of large tables. 

- Your final PDF should be 18 pages. 

- You can insert needed page breaks as illustrated below

- Make sure you do not "print" the data. If you change the data, make sure you store with a structure like: newname<-modification(olddata). If you only type modification(olddata), R will display the data rather than saving your modifications


3) Upload your assignment PDF to gradescope.

\pagebreak

# Set up and opening the data
## Question 1.1: 
**Load the `basic.dta` file like you did for PS1 and call all the packages you will be using with the `library` function. The packages you will need are `haven`, `lfe`,  `dplyr`,  and `stargazer`.**


```{r,warning = FALSE, message = FALSE}
library("haven")
library("lfe")
library("dplyr")
library("stargazer")#call packages
```


```{r}
dataovb <- read_dta("C:\\Users\\Nachu\\Downloads\\basic.dta")#load data
```



\pagebreak
# Cleaning the data
## Question 2.1: 
**Like in PS1, restrict your data to observations from the 105th congress and keep only the variables listed in the table below. Here too, make sure your final dataset is a data frame. **


| Name     |Description                                                             |
|----------|------------------------------------------------------------------------|
|aauw	     |AAUW score                                                              | 
|totchi	   |Total number of children                                                |
|ngirls	   |Number of daughters                                                     |
|party	   |Political party. Democrats if 1, Republicans if 2, and Independent if 3.|
|female	   |Female dummy variable                                                   |
|white	   |White dummy variable                                                    |
|srvlng	   |Years of service                                                        | 
|age	     |Age                                                                     |
|demvote	 |State democratic vote share in most recent presidential election        |
|rgroup    |religious group                                                         |
|region    |region                                                                  |
|name      |representative's name                                                                 |

**You can find the detailed description of each variable in the original paper. The main variable in this analysis is `AAUW`, a score created by the American Association of University Women (AAUW). For each congress, AAUW selects pieces of legislation in the areas of education, equality, and reproductive rights. The AAUW keeps track of how each legislator voted on these pieces of legislation and whether their vote aligned with the AAUW’s position. The legislator’s score is equal to the proportion of these votes made in agreement with the AAUW.**

Hint: You can use the `filter` and `select` functions we discussed in Slides2_IntroR.


```{r}
#clean data - filter 105th congress, select reqd col
dataovb <- dataovb %>% filter(congress == 105) %>% 
  select(aauw,totchi,ngirls,party,
         female,white,srvlng,age,demvote,rgroup,region,name) 

dataovb <- as.data.frame(dataovb) #convert to dataframe
```





\pagebreak
# Analysis
## Question 3.1: 
**Estimate the following linear regression models using the `felm` command (part of the lfe package). Report your regression results in a formatted table using  `stargazer`. Report robust standard errors in your table.** 


$$
\begin{aligned}
 \text{Model 1: }aauw_i&=\beta_0+\beta_1ngirls_i+\epsilon_i\\
 \text{Model 2: }aauw_i&=\beta_0+\beta_1ngirls_i+\beta_2totchi+\epsilon_i\\
\end{aligned}
$$
Hints: If you want RMarkdown to display your outputted table, include the code `results = "asis"` in the chunk header. This is true for all chunks that output a formatted table. In the stargazer command, you will want to specify the format of the table by including the code `type="latex"` for a pdf output. If you have trouble knitting to PDF, try installing MikTex (https://miktex.org/download). You will need to specify `se = list(reg1$rse, reg2$rse)` in stargazer, like the example in Slides3_RegressionReview.

```{r, results = "asis"}
Model1ovb <- felm(aauw ~ ngirls, data = dataovb)
Model2ovb <- felm(aauw ~ ngirls + totchi, data = dataovb)

stargazer(Model1ovb, Model2ovb,
          se = list(Model1ovb$rse, Model2ovb$rse),
          type = "latex",
          title = "Regression Results - 3.1",
          column.labels = c("Model 1","Model 2"),
          covariate.labels = c("Number of daughters", 
                               "Total Children" )) 


```



\pagebreak
## Question 3.2: 

**Compare the estimates of $\beta_1$ across the two specifications. Why does our estimate of $\beta_1$ changes so much? Which control variable is particularly important and why? **

**Answer:**
Beta 1 changes because of the introduction of control variable - total children. Having more daughters is dependent on having more total children. Without that control, there is a negative correlation between number of daughters and AAUW score. Having more children is generally associated with conservative views, and the isolated effect of daughters on AAUW score is reflected only after the introduction of control variable. Having total children as control variable is particularly important to separate the effects on AAUW score - the negative effect of having more children, and the positive effect of having daughters.


\pagebreak
## Question 3.3: 
**Consider the second specification which controls for $totchi_i$. Conditional on the number of children, do you think $ngirls_i$ is plausibly exogenous? What is the identifying assumption, i.e. the assumption that must hold for $\beta_1$ to be interpreted as a causal estimate? What evidence does Washington give to support this assumption?**

**Answer:**
Yes ngirls is plausibly exogenous. The identifying assumption is that the gender of the child is determined by nature - factors not under the congress person's control. The number of daughters being a random variable must be true for beta1 to be estimated as a causal estimate. Washington provides evidence that Congress people are not following fertility stopping rules that show preference to one gender over the other, and the gender of the child is truly random.

\pagebreak
# Fixed Effects:
## Question 4.1 (2 pages): 
**Equation 1 from Washington's paper is a little bit different from the equations you have estimated so far. Estimate the three models specified below (where $\gamma_i$ is a fixed effect for the number of children). Present your results in a table. Use robust standard errors.**

$$
\begin{aligned}
  \text{Model 1: }aauw_i&=\beta_0+\beta_1ngirls_i+\beta_2totchi+\epsilon_i\\
   \text{Model 2: }aauw_i&=\beta_0+\beta_1ngirls_i+\beta_2chi1+...+\beta_{10}chi10 +\epsilon_i\\
    \text{Model 3: } aauw_i&=\beta_0+\beta_1ngirls_i+\gamma_i+\epsilon_i\\
\end{aligned}
$$

Hints: 

- you will need to generate the dummy variables for the second equation or code it as `factor(totchi)`. 

- For the third equation, the `felm` function allows you to specify fixed effects as we saw in class. 

**Code:**
```{r, results = "asis"}
Model2ovb <- felm(aauw ~ ngirls + totchi, data = dataovb)

Model3ovb <- felm(aauw ~ ngirls + factor(totchi), data = dataovb)

Model4ovb <- felm(aauw ~ ngirls|totchi, data = dataovb)

stargazer(Model2ovb, Model3ovb,Model4ovb,
          se = list(Model1ovb$rse, Model2ovb$rse),
          type = "latex",
          title = "Regression Results - 4.1",
          column.labels = c("Model 1","Model 2", "Model3"),
          covariate.labels = c("Number of daughters", 
                               "Total Children" )) 
```



\pagebreak
## Question 4.2: 
**Explain the difference between the three models.**

**Answer:**
Model 1 regresses the number of girls and total children on the AAUW score. 
Model 2 regresses the number of girls and the effect of each additional child with 0 children as a dummy variable, on the AAUW score. 
Model 3 regresses the number of girls on the AAUW score, keeping total number of children as the fixed effect. 

The Beta co-efficient for ngirls for all models are similar because, gender is a random variable, and the beta for total children and number of daughters effectively reflect the isolated effects of having an additional daughter, and an additional child. This is true regardless of the model - controlling for it as an additional variable, or factorizing total children variable, or having total children as a fixed effect.  The number of children, generally, has a negative effect on the score. The Beta of having an additional daughter has a positive effect on AAUW score and is almost the same across all 3 models. 


\pagebreak
## Question 4.3 (2 pages): 
**Reproduce the EXACT results presented in column 2 of table 2 from Washington's paper. To do this you will need to first build three variables: $age^2$ and $srvlng^2$ and $repub_i$, an indicator set to 1 if the representative is republican and 0 otherwise.  Then estimate the following specification, where $\gamma_i$ is a fixed effect for total children, $\phi_i$ is a fixed effect for religious group, and $\lambda_i$ is a fixed effect for region:**

$$
\text{Model A: }aauw_i=\beta_0+\beta_1ngirls_i+female_i+white_i+repub_i+age_i+age^2_i+srvlng_i+srvlng^2_i+demvote_i+\gamma_i+\phi_i+\lambda_i+\epsilon_i
$$
Hint: 

- Because you are using a fixed effect for religion, stargazer will not print the religion coefficients. That is ok, those coefficients are still being estimated in the background of the R program. 

- Do not use robust standard errors as she does not in this table. 

**Code:**
```{r, results = "asis"}
dataovb <- dataovb %>% mutate(age2 = age^2, srvlng2 = srvlng^2, repub_i = ifelse(party == 2, 1, 0)) 
# add 3 new variables - age square, years of service square and dummy variable for repub

ModelAovb <- felm(aauw~ngirls+female+white+repub_i+age+age2+
                    srvlng+srvlng2+demvote |totchi+rgroup+region, 
                  data = dataovb)


stargazer(ModelAovb,
          type = "latex",
          title = "Regression Results - 4.3",
          column.labels = c("Model A"),
          covariate.labels = c("Number of daughters", 
                               "Female",
                               "White","Republican",
                               "Age","Age squared","Serv length","Serv length squared",
                               "Democratic vote"))
          
                

```



\pagebreak
## Question 4.4: 
**Explain what the region fixed effects are controlling for.**

**Answer:**
Region fixed effects control for variations in voting patterns across different regions. The effect of having an additional daughter is isolated from the effect of the congress person belonging to a particular region. Region is held constant, and just the effect of an additional daughter is measured. 

\pagebreak
## Question 4.5: 
**Reload the data and this time keep observations from all of the four congresses. Add the three variables you built for question 4.3 to this data set**

**Code:**
```{r}
dataovb_reload <- read_dta("C:\\Users\\Nachu\\Downloads\\basic.dta")#load data
dataovb_reload <- dataovb_reload %>% mutate(age2 = age^2, srvlng2 = srvlng^2, 
                  repub_i = ifelse(party == 2, 1, 0)) 
#mutate additional variables
```

\pagebreak
## Question 4.6 (2 pages): 
**Because we have data for four congress sessions, we may be able to see how an individual congress person's voting patterns change as the number of daughters they have changes. Estimate model A with the addition of `congress` and `name` fixed effects. Present your results in a table.** 

**Code:**
```{r, results = "asis"}
ModelAovb_reload <- felm(aauw~ngirls+female+white+repub_i+age+age2+srvlng+srvlng2
                         +demvote |totchi+rgroup+region+congress+name, data = dataovb_reload)


stargazer(ModelAovb_reload,
          type = "latex",
          title = "Regression Results - 4.6",
          column.labels = c("Model A"),
          covariate.labels = c("Number of daughters", 
                               "Female",
                               "White","Republican",
                               "Age","Age squared","Serv length",
                               "Serv length squared",
                               "Democratic vote"))
                
```


\pagebreak
## Question 4.7: 
**How does this estimate compare to your estimate in question 4.3? Why are the standard errors so much bigger? Why doesn't Washington use this approach in her paper?** 

**Answer:**
The beta of  having an additional daughter is lower than the previous model. The standard errors are much higher because the number of congress men with change in ngirls becomes lower.Washington does not use this approach because of low sample size and higher standard errors. 

\pagebreak
## Question 4.8:
**Why are you not able to generate a coefficient for $female_i$ or $white_i$?** 

**Answer:**
There is no co-efficient for female or white because gender and race do not change across the congress session, and therefore have no effect on the AAUW score across congress sessions.

\pagebreak
## Question 4.9: 
**You are able to generate an estimate for $repub_i$. What does this imply?** 

**Answer:**
This implies that a few congress members switched parties across congress sessions and the effect of the switch on the AAUW score is reflected in the Beta estimate. 
`




